% kubernetes-vulns.tex
% Thesis section: Extending MAD Goat to the Kubernetes Layer
% Compiled from thesis/kubernetes-vulns/ with:
%   pdflatex -shell-escape test-main.tex

\section{Extending the MAD Goat Benchmark to the Kubernetes Layer}
\label{sec:k8s-vulns}

\subsection{Motivation and Context}
\label{sec:k8s-motivation}

The MAD~Goat benchmark~\cite{mad-goat-repo} was originally designed to surface web-application and microservice-level vulnerabilities in containerised workloads.
As Kubernetes adoption has grown, the attack surface has shifted from individual containers to the orchestration layer itself: cluster RBAC, admission controllers, network policies, kubelet configuration, and audit infrastructure all introduce exploitable misconfigurations that are invisible to application-layer scanners~\cite{shamim-2020,bose-2023}.
The OWASP Kubernetes Top~10 2025 edition~\cite{owasp-k8s-2025} codifies ten categories of such weaknesses, updating and renumbering the 2022 list~\cite{owasp-k8s-2022} (Table~\ref{tab:owasp-crosswalk}).
This section extends MAD~Goat to the Kubernetes layer by injecting all ten categories into the benchmark's Helm charts and cluster bootstrap configuration, providing a reproducible, toggle-controlled environment for evaluating both active exploits and static analysis tools.

\input{tables/crosswalk-2022-2025}

\subsection{Methodology}
\label{sec:k8s-methodology}

Each vulnerability entry is controlled by a boolean flag in the chart's \texttt{values.yaml} (default: \texttt{true} = vulnerable).
Chart-level entries (\texttt{scope: chart}) live in \texttt{helm/madgoat} or \texttt{helm/madgoat-infra} and can be toggled with \texttt{helm upgrade}.
Cluster-level entries (\texttt{scope: cluster}) are applied via \texttt{kubectl apply -k k8s/insecure-defaults/} and require a full Minikube restart to remediate; the \texttt{make reset-safe} target automates this.
Kubelet flags (K07) are passed as \texttt{minikube start -{}-extra-config} arguments and are therefore not Helm-controlled.
Two entries (K08) are conceptual only: the IMDS credential-theft attack surface does not exist in a local Minikube cluster and is documented with references to cloud-vendor behaviour~\cite{aws-imds,azure-imds}.

For each \emph{full}-class entry the benchmark provides:
\begin{itemize}
  \item A \texttt{values.yaml} boolean flag that injects or suppresses the misconfiguration.
  \item A verification script in \texttt{scripts/exploits/} that exits~0 (PASS) when the vulnerable condition is confirmed and exits~1 (FAIL/safe) otherwise.
  \item Code snippets (Listings below) illustrating the vulnerable and remediated manifests.
\end{itemize}

The ground-truth inventory is maintained in \texttt{thesis/kubernetes-vulns/data/vulnerabilities.yaml}; Tables~\ref{tab:vuln-inventory} and~\ref{tab:scanner-coverage} are generated from it automatically by \texttt{scripts/gen-tables.py}~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\input{tables/vuln-inventory}

\subsection{Vulnerability Catalogue}
\label{sec:k8s-catalogue}

% ─── K01 ────────────────────────────────────────────────────────────────────
\subsubsection{K01 --- Insecure Workload Configurations (2022: K01)}
\label{sec:k01}

K01 covers pod-level security defaults that widen the blast radius of any container compromise.
MAD~Goat injects four orthogonal K01 sub-entries across both charts via independent boolean flags.

\paragraph{K01-a: Run-as-root containers (CWE-250, CIS~5.2.6).}\label{sec:k01-runAsRoot}
When no \texttt{securityContext} is applied, Kubernetes allows containers to execute as \textsc{uid}~0.
A successful container escape from a root process gains equivalent privileges on the host filesystem.
The flag \texttt{vulnerabilities.k01\_runAsRoot=true} (default) suppresses the security context on every pod in both charts; setting it to \texttt{false} injects \texttt{runAsNonRoot: true}, \texttt{runAsUser: 10001}, and \texttt{allowPrivilegeEscalation: false}~\cite{nsa-cisa-k8s-2022,cis-k8s-1.9}.
Static scanners detect this pattern: KICS rule \texttt{cf34805e}, Trivy \texttt{KSV012}, KubeLinter \texttt{run-as-non-root}, Kubescape \texttt{C-0013}~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-runAsRoot-vuln.yaml}
\caption{K01-a vulnerable: \texttt{securityContext} absent --- container runs as root.}\label{lst:k01-runAsRoot-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-runAsRoot-safe.yaml}
\caption{K01-a remediated: \texttt{runAsNonRoot} and non-zero \texttt{runAsUser} enforced.}\label{lst:k01-runAsRoot-safe}
\end{listing}

\paragraph{K01-b: Writable root filesystem (CWE-276).}\label{sec:k01-readOnlyRootFs}
Without \texttt{readOnlyRootFilesystem: true}, an attacker with code execution can overwrite container binaries or drop persistence payloads.
The flag \texttt{vulnerabilities.k01\_readOnlyRootFs} controls this field on all pods in both charts.
Trivy check \texttt{KSV014}, KICS rule \texttt{a9c2f49d}, KubeLinter \texttt{read-only-root-fs}, and Kubescape \texttt{C-0017} all flag its absence~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-readOnlyRootFs-vuln.yaml}
\caption{K01-b vulnerable: empty \texttt{securityContext} leaves filesystem writable.}\label{lst:k01-readOnlyRootFs-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-readOnlyRootFs-safe.yaml}
\caption{K01-b remediated: \texttt{readOnlyRootFilesystem: true} applied.}\label{lst:k01-readOnlyRootFs-safe}
\end{listing}

\paragraph{K01-c: Privileged containers (CWE-250, CIS~5.2.2).}\label{sec:k01-privileged}
Setting \texttt{privileged: true} disables all Linux namespace isolation between the container and the host kernel~\cite{nist-sp800-190}.
MAD~Goat applies this to three services in \texttt{helm/madgoat} (\texttt{mad4shell-unsafe}, \texttt{profile}, \texttt{webapp}) that model intentionally dangerous images.
Trivy check \texttt{KSV017}, KICS rule \texttt{dd29336b}, KubeLinter \texttt{privileged-container}, and Kubescape \texttt{C-0016} detect this condition~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-privileged-vuln.yaml}
\caption{K01-c vulnerable: \texttt{privileged: true} on \texttt{mad4shell-unsafe}.}\label{lst:k01-privileged-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-privileged-safe.yaml}
\caption{K01-c remediated: \texttt{privileged} absent (defaults to \texttt{false}).}\label{lst:k01-privileged-safe}
\end{listing}

\paragraph{K01-d: Unbounded resource usage (CWE-400).}\label{sec:k01-unboundedResources}
A container without \texttt{resources.limits} can consume unlimited CPU and memory, enabling denial-of-service against co-located workloads~\cite{nsa-cisa-k8s-2022}.
The flag \texttt{vulnerabilities.k01\_unboundedResources} controls limit injection on the \texttt{lesson} pod in \texttt{helm/madgoat}.
Trivy checks \texttt{KSV011}/\texttt{KSV018}, KICS rule \texttt{b14d1bc4}, and Kubescape \texttt{C-0009} detect this~\cite{trivy-k8s,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-unboundedResources-vuln.yaml}
\caption{K01-d vulnerable: \texttt{resources} field absent on \texttt{lesson} pod.}\label{lst:k01-unboundedResources-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k01-unboundedResources-safe.yaml}
\caption{K01-d remediated: CPU and memory \texttt{requests}/\texttt{limits} applied.}\label{lst:k01-unboundedResources-safe}
\end{listing}

% ─── K02 ────────────────────────────────────────────────────────────────────
\subsubsection{K02 --- Overly Permissive RBAC (2022: K03)}
\label{sec:k02}

Over-privileged RBAC bindings are the primary lateral-movement path in a compromised Kubernetes cluster~\cite{k8s-rbac-docs,shamim-2020}.
MAD~Goat injects two complementary RBAC misconfigurations into \texttt{helm/madgoat-infra}.

\paragraph{K02-a: Default ServiceAccount bound to \texttt{cluster-admin} (CWE-269, CIS~5.1.1).}\label{sec:k02-clusterAdminBinding}
Binding the \texttt{default} ServiceAccount to the built-in \texttt{cluster-admin} ClusterRole means any pod in the \texttt{default} namespace that mounts its service-account token can perform any action on any resource cluster-wide.
The flag \texttt{vulnerabilities.k02\_clusterAdminBinding=true} emits the ClusterRoleBinding in Listing~\ref{lst:k02-clusterAdminBinding-vuln}; setting it to \texttt{false} suppresses the resource entirely.
Detected by: Trivy \texttt{KSV111}, KICS rule \texttt{249328b8}, KubeLinter \texttt{cluster-admin-role-binding}, Kubescape \texttt{C-0035}~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k02-clusterAdminBinding-vuln.yaml}
\caption{K02-a vulnerable: \texttt{default} SA bound to \texttt{cluster-admin}.}\label{lst:k02-clusterAdminBinding-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k02-clusterAdminBinding-safe.yaml}
\caption{K02-a remediated: ClusterRoleBinding absent.}\label{lst:k02-clusterAdminBinding-safe}
\end{listing}

\paragraph{K02-b: Role with \texttt{list}/\texttt{watch} on Secrets (CWE-732, CIS~5.1.3).}\label{sec:k02-secretsListWatch}
A Role granting \texttt{list} and \texttt{watch} verbs on the \texttt{secrets} resource enables any pod bound to that Role to enumerate and continuously watch all Secret objects in the namespace, effectively providing a bulk credential dump~\cite{k8s-rbac-docs}.
Trivy \texttt{KSV041}, KICS rule \texttt{b7bca5c4}, KubeLinter \texttt{access-to-secrets}, and Kubescape \texttt{C-0015} detect this~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k02-secretsListWatch-vuln.yaml}
\caption{K02-b vulnerable: Role permits \texttt{list}+\texttt{watch} on all Secrets.}\label{lst:k02-secretsListWatch-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k02-secretsListWatch-safe.yaml}
\caption{K02-b remediated: Role absent.}\label{lst:k02-secretsListWatch-safe}
\end{listing}

% ─── K03 ────────────────────────────────────────────────────────────────────
\subsubsection{K03 --- Secrets Management Failures (2022: K08)}
\label{sec:k03}

Kubernetes \texttt{ConfigMap} objects are stored unencrypted in \texttt{etcd} and are readable by any principal with \texttt{get} access to the namespace~\cite{k8s-secrets-docs}.
Both K03 entries were \emph{already present} in MAD~Goat before this extension (\texttt{presentPreInjection: true}); the flags retrofit a toggle-controlled migration path.

\paragraph{K03-a: Credentials in plaintext ConfigMap (CWE-522, CIS~5.4.1).}\label{sec:k03-plaintextConfigMap}
The \texttt{mad-db-config} ConfigMap stores PostgreSQL and MongoDB root credentials in plaintext.
Any pod that can execute \texttt{kubectl get configmap mad-db-config} or that mounts the ConfigMap as an environment source immediately obtains database credentials.
The flag \texttt{vulnerabilities.k03\_plaintextConfigMapSecrets=false} migrates these values to a \texttt{Secret}~\cite{k8s-secrets-docs,nsa-cisa-k8s-2022}.
Detected by: Trivy \texttt{AVD-KSV-0109}, KICS rule \texttt{3d658f8b}, Kubescape \texttt{C-0207}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k03-plaintextConfigMap-vuln.yaml}
\caption{K03-a vulnerable: database credentials in a \texttt{ConfigMap}.}\label{lst:k03-plaintextConfigMap-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k03-plaintextConfigMap-safe.yaml}
\caption{K03-a remediated: credentials moved to a \texttt{Secret}.}\label{lst:k03-plaintextConfigMap-safe}
\end{listing}

\paragraph{K03-b: JWT signing key in ConfigMap environment variable (CWE-522, CIS~5.4.2).}\label{sec:k03-jwtKeyInEnv}
The authentication service reads its HMAC JWT signing key from the \texttt{JWT\_SECRET} field of \texttt{mad-auth-config}.
An attacker with ConfigMap read access can forge arbitrary JWT tokens and impersonate any user.
Setting \texttt{vulnerabilities.k03\_jwtKeyInEnv=false} moves the key to a \texttt{Secret} referenced via \texttt{secretKeyRef}.
Detected by: Trivy \texttt{AVD-KSV-0109}, KICS rule \texttt{b9c83569}, Kubescape \texttt{C-0066}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k03-jwtKeyInEnv-vuln.yaml}
\caption{K03-b vulnerable: JWT key in plaintext \texttt{ConfigMap}.}\label{lst:k03-jwtKeyInEnv-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k03-jwtKeyInEnv-safe.yaml}
\caption{K03-b remediated: JWT key moved to a \texttt{Secret}.}\label{lst:k03-jwtKeyInEnv-safe}
\end{listing}

% ─── K04 ────────────────────────────────────────────────────────────────────
\subsubsection{K04 --- Lack of Centralised Policy Enforcement (2022: K04)}
\label{sec:k04}

Admission controllers are the last line of defence before a pod manifest reaches the scheduler.
Without enforcement-mode policies, all K01 misconfigurations can be deployed regardless of image scanner findings~\cite{k8s-psa-docs,kyverno-docs}.
Both K04 entries are cluster-level (\texttt{scope: cluster}) and are applied via \texttt{kubectl apply -k k8s/insecure-defaults/}.

\paragraph{K04-a: Namespace without Pod Security Admission enforcement (CWE-693, CIS~1.2.9).}\label{sec:k04-noPSA}
The \texttt{default} namespace is labelled with \texttt{pod-security.kubernetes.io/enforce: privileged}, which permits any pod security profile including \texttt{privileged: true}~\cite{k8s-psa-docs}.
This was already the Minikube default (\texttt{presentPreInjection: true}); the benchmark makes it explicit.
The exploit script \texttt{scripts/exploits/k04-noPSA.sh} verifies the label value.
Detected by: Trivy \texttt{KSV0010}, KICS rule \texttt{ce30e584}, Kubescape \texttt{C-0122}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k04-noPSA-vuln.yaml}
\caption{K04-a vulnerable: \texttt{default} namespace enforces \texttt{privileged} PSA profile.}\label{lst:k04-noPSA-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k04-noPSA-safe.yaml}
\caption{K04-a remediated: enforce \texttt{baseline}, audit \texttt{restricted}.}\label{lst:k04-noPSA-safe}
\end{listing}

\paragraph{K04-b: Kyverno installed with no-op \texttt{Audit} rules (CWE-693, CIS~1.2.12).}\label{sec:k04-weakKyverno}
Installing Kyverno with \texttt{validationFailureAction: Audit} creates the false impression of policy enforcement: violations are written to the Kyverno event log but non-compliant pods are admitted.
An operator inspecting \texttt{kubectl get clusterpolicy} would see active policies without realising they are non-blocking.
The benchmark applies the weak ClusterPolicy to demonstrate this false-positive governance pattern~\cite{kyverno-docs}.
Kubescape \texttt{C-0123} detects Audit-mode policies~\cite{kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k04-weakKyverno-vuln.yaml}
\caption{K04-b vulnerable: \texttt{validationFailureAction: Audit} --- violations logged but permitted.}\label{lst:k04-weakKyverno-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k04-weakKyverno-safe.yaml}
\caption{K04-b remediated: \texttt{validationFailureAction: Enforce} --- violations blocked.}\label{lst:k04-weakKyverno-safe}
\end{listing}

% ─── K05 ────────────────────────────────────────────────────────────────────
\subsubsection{K05 --- Inadequate Network Segmentation (2022: K07)}
\label{sec:k05}

Without \texttt{NetworkPolicy} objects the Kubernetes virtual network is fully flat: every pod can reach every other pod and any external endpoint~\cite{k8s-netpol-docs}.
MAD~Goat injects three complementary K05 sub-entries into \texttt{helm/madgoat-infra} that were all \emph{already present} in the pre-injection baseline (\texttt{presentPreInjection: true}).

\paragraph{K05-a: No default-deny \texttt{NetworkPolicy} (CWE-923, CIS~5.3.2).}\label{sec:k05-noDefaultDeny}
The absence of a deny-all baseline policy means that compromising any pod provides free lateral movement to every other pod without any network-layer control.
Setting \texttt{vulnerabilities.k05\_noDefaultDeny=false} adds the \texttt{default-deny-all} policy in Listing~\ref{lst:k05-noDefaultDeny-safe}.
Detected by: Trivy \texttt{KSV038}, KICS rule \texttt{03aabc8c}, KubeLinter \texttt{non-isolated-pod}, Kubescape \texttt{C-0205}~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-noDefaultDeny-vuln.yaml}
\caption{K05-a vulnerable: no default-deny \texttt{NetworkPolicy} present.}\label{lst:k05-noDefaultDeny-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-noDefaultDeny-safe.yaml}
\caption{K05-a remediated: \texttt{default-deny-all} blocks all uninstructed traffic.}\label{lst:k05-noDefaultDeny-safe}
\end{listing}

\paragraph{K05-b: Overly permissive label-based \texttt{NetworkPolicy} (CWE-923, CIS~5.3.1).}\label{sec:k05-permissiveNetpol}
The pre-existing \texttt{mad-network-policy} uses a broad \texttt{podSelector} matching the \texttt{io.mad.service} label, which is shared by all MAD~Goat pods.
This effectively grants every service access to every other service, defeating the intent of the policy.
Setting \texttt{vulnerabilities.k05\_permissiveNetpol=false} replaces it with explicit per-service ingress rules.
Detected by: KICS rule \texttt{85ab1c5b}, KubeLinter \texttt{dangling-networkpolicy}, Kubescape \texttt{C-0206}~\cite{kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-permissiveNetpol-vuln.yaml}
\caption{K05-b vulnerable: label-based policy permits all MAD~Goat pods to communicate.}\label{lst:k05-permissiveNetpol-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-permissiveNetpol-safe.yaml}
\caption{K05-b remediated: explicit per-service ingress rule.}\label{lst:k05-permissiveNetpol-safe}
\end{listing}

\paragraph{K05-c: Missing egress controls (CWE-923, CIS~5.3.1).}\label{sec:k05-noEgressControls}
With no egress \texttt{NetworkPolicy}, any compromised pod can initiate outbound connections to arbitrary internet endpoints, enabling command-and-control callbacks, DNS exfiltration, or cloud-metadata queries~\cite{aws-imds}.
Setting \texttt{vulnerabilities.k05\_noEgressControls=false} adds an allowlist-based egress policy restricting outbound traffic to DNS and internal database services.
Detected by: Trivy \texttt{KSV038}, KICS rule \texttt{0401f71b}, KubeLinter \texttt{dangling-networkpolicypeer-podselector}, Kubescape \texttt{C-0054}~\cite{trivy-k8s,kics-tool,kubelinter,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-noEgressControls-vuln.yaml}
\caption{K05-c vulnerable: no egress \texttt{NetworkPolicy}.}\label{lst:k05-noEgressControls-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k05-noEgressControls-safe.yaml}
\caption{K05-c remediated: egress allowlist restricts outbound to DNS and internal services.}\label{lst:k05-noEgressControls-safe}
\end{listing}

% ─── K06 ────────────────────────────────────────────────────────────────────
\subsubsection{K06 --- Broken Authentication Mechanisms (new in 2025)}
\label{sec:k06}

K06 is a new category in the 2025 edition with no direct 2022 equivalent.
It covers administrative interfaces and management planes that are exposed without authentication controls.
MAD~Goat injects two sub-entries into \texttt{helm/madgoat-infra}; both were present in the original deployment (\texttt{presentPreInjection: true}).

\paragraph{K06-a: Traefik dashboard publicly exposed (CWE-284).}\label{sec:k06-traefikDashboard}
Traefik ships with a dashboard and API at \texttt{/dashboard} and \texttt{/api} that expose the full routing configuration, middleware chain, and service health status.
The \texttt{IngressRoute} in Listing~\ref{lst:k06-traefikDashboard-vuln} routes external requests to the \texttt{api@internal} service with no authentication middleware~\cite{traefik-docs}.
An attacker can enumerate all routes, backends, and TLS certificates.
Setting \texttt{vulnerabilities.k06\_traefikDashboardExposed=false} suppresses the \texttt{IngressRoute} entirely.
No KICS, Trivy, KubeLinter, or Kubescape rules cover this Traefik-specific CRD pattern --- making it an example of an application-layer misconfiguration invisible to generic manifest scanners.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k06-traefikDashboard-vuln.yaml}
\caption{K06-a vulnerable: Traefik dashboard \texttt{IngressRoute} without authentication.}\label{lst:k06-traefikDashboard-vuln}
\end{listing}

\paragraph{K06-b: RabbitMQ management UI publicly exposed (CWE-284).}\label{sec:k06-rabbitmqMgmt}
RabbitMQ's management plugin exposes an HTTP API and web UI on port~15672 with default credentials \texttt{guest}/\texttt{guest}.
The \texttt{IngressRoute} in Listing~\ref{lst:k06-rabbitmqMgmt-vuln} routes \texttt{rabbitmq.localhost} directly to this port.
An unauthenticated attacker can enumerate queues, publish or consume messages, and obtain AMQP connection credentials.
Setting \texttt{vulnerabilities.k06\_rabbitmqMgmtExposed=false} removes the route; the management port remains internal-only.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k06-rabbitmqMgmt-vuln.yaml}
\caption{K06-b vulnerable: RabbitMQ management UI exposed on \texttt{rabbitmq.localhost}.}\label{lst:k06-rabbitmqMgmt-vuln}
\end{listing}

% ─── K07 ────────────────────────────────────────────────────────────────────
\subsubsection{K07 --- Misconfigured Cluster Components (2022: K09)}
\label{sec:k07}

The kubelet is the per-node agent that executes pod manifests; its HTTP API exposes pod logs, \texttt{exec} endpoints, and node metrics.
K07 kubelet entries are cluster-level and not Helm-controlled: they require \texttt{minikube start -{}-extra-config} flags and are provisioned by \texttt{make reset-vulnerable}~\cite{k8s-kubelet-docs}.

\paragraph{K07-a: \texttt{kubelet} anonymous authentication enabled (CWE-306, CIS~1.2.1).}\label{sec:k07-kubeletAnonymousAuth}
Setting \texttt{authentication.anonymous.enabled=true} instructs the kubelet to accept unauthenticated requests as the \texttt{system:anonymous} user.
An attacker with network access to port~10250 can retrieve pod logs, environment variables, and container filesystem listings without credentials.
The exploit script \texttt{scripts/exploits/k07-kubeletAnonymousAuth.sh} confirms the condition by querying the kubelet \texttt{/healthz} endpoint without a token.
Detected by: Trivy \texttt{KCV0001}, KICS rule \texttt{1de5cc51}, Kubescape \texttt{C-0113}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k07-kubelet-vuln.yaml}
\caption{K07 vulnerable: kubelet with \texttt{anonymous-auth=true} and \texttt{AlwaysAllow} authorisation.}\label{lst:k07-kubelet-vuln}
\end{listing}

\paragraph{K07-b: \texttt{kubelet} authorisation mode set to \texttt{AlwaysAllow} (CWE-306, CIS~1.2.6).}\label{sec:k07-kubeletAlwaysAllow}
Even when authentication is enabled, \texttt{authorization.mode=AlwaysAllow} grants every authenticated identity full access to every kubelet API endpoint, including \texttt{/run} (arbitrary command execution on the node).
This is combined with K07-a in \texttt{make reset-vulnerable} to produce a fully open kubelet.
Detected by: Trivy \texttt{KCV0007}, KICS rule \texttt{f1f4d8da}, Kubescape \texttt{C-0120}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k07-kubelet-safe.yaml}
\caption{K07 remediated: webhook authentication and authorisation enforced by \texttt{make reset-safe}.}\label{lst:k07-kubelet-safe}
\end{listing}

% ─── K08 ────────────────────────────────────────────────────────────────────
\subsubsection{K08 --- Supply Chain Vulnerabilities (2022: K02)}
\label{sec:k08}

K08 covers two attack paths that exploit trust in the supply chain of a running cluster.
Both entries are classified \texttt{class: conceptual} because their prerequisites (cloud-provider IMDS, managed Kubernetes node IAM roles) do not exist in a local Minikube cluster.
They are documented here as design patterns with cloud-environment references~\cite{aws-imds,azure-imds}.

\paragraph{K08-a: IMDS credential theft via SSRF (conceptual).}\label{sec:k08-imdsCredentialTheft}
In cloud-hosted Kubernetes clusters (EKS, AKS, GKE), the Instance Metadata Service (IMDS) is reachable at \texttt{169.254.169.254} from within any pod unless explicitly blocked.
An SSRF vulnerability in any MAD~Goat service can be chained to a request to \texttt{http://169.254.169.254/latest/meta-data/iam/security-credentials/} to obtain temporary AWS IAM credentials for the node's instance profile~\cite{aws-imds}.
The absence of K05-c egress controls (Section~\ref{sec:k05-noEgressControls}) enables this attack path.
Mitigation requires IMDSv2 enforcement (\texttt{HttpTokens: required}) and an egress \texttt{NetworkPolicy} blocking \texttt{169.254.0.0/16}.

\paragraph{K08-b: Service-account token projection to cloud (conceptual).}\label{sec:k08-saTokenToCloud}
Kubernetes service-account tokens projected into pods via the \texttt{BoundServiceAccountTokenVolume} feature can be exchanged for cloud-provider credentials using OIDC workload identity federation.
Combined with the K09-b shared default ServiceAccount (Section~\ref{sec:k09-sharedServiceAccount}), a compromised pod can present its token to the cloud IAM endpoint and assume the node's workload identity~\cite{azure-imds}.
Mitigation requires per-workload ServiceAccounts with explicit OIDC audience bindings and shortest-possible token TTLs.

% ─── K09 ────────────────────────────────────────────────────────────────────
\subsubsection{K09 --- Misconfigured Authentication Defaults (2022: K06)}
\label{sec:k09}

K09 groups authentication defaults in cluster components and workloads that allow unauthenticated or weakly authenticated access.
MAD~Goat injects two K09 sub-entries into \texttt{helm/madgoat-infra} and both charts respectively.

\paragraph{K09-a: Keycloak default admin credentials and \texttt{start-dev} mode (CWE-200).}\label{sec:k09-anonymousKeycloakBootstrap}
Keycloak's \texttt{start-dev} command disables HTTPS, uses an embedded H2 database, and starts with default credentials \texttt{admin}/\texttt{admin}~\cite{keycloak-docs}.
This was already the MAD~Goat deployment configuration (\texttt{presentPreInjection: true}).
The flag \texttt{vulnerabilities.k09\_anonymousKeycloakBootstrap=false} switches the Keycloak container command to \texttt{start} (production mode), enforcing HTTPS and preventing unauthenticated admin console access.
No generic scanner rules exist for application-level credential defaults; the exploit script \texttt{scripts/exploits/k09-anonymousKeycloakBootstrap.sh} verifies HTTP~200 from the admin console endpoint without credentials.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k09-keycloak-vuln.yaml}
\caption{K09-a vulnerable: Keycloak in \texttt{start-dev} with default \texttt{admin}/\texttt{admin}.}\label{lst:k09-keycloak-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k09-keycloak-safe.yaml}
\caption{K09-a remediated: Keycloak in production \texttt{start} mode.}\label{lst:k09-keycloak-safe}
\end{listing}

\paragraph{K09-b: All pods share the \texttt{default} ServiceAccount (CWE-200).}\label{sec:k09-sharedServiceAccount}
When no explicit \texttt{serviceAccountName} is set, Kubernetes assigns the \texttt{default} ServiceAccount, which auto-mounts a JWT token into every pod at \texttt{/var/run/secrets/kubernetes.io/serviceaccount/token}~\cite{k8s-rbac-docs}.
Combined with K02-a (cluster-admin binding on \texttt{default}), any compromised pod can use this token to control the cluster.
The flag \texttt{vulnerabilities.k09\_sharedServiceAccount=false} creates a dedicated per-service \texttt{ServiceAccount} with \texttt{automountServiceAccountToken: false} for every service in both charts.
Kubescape \texttt{C-0183} detects automounted tokens~\cite{kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k09-sharedSA-vuln.yaml}
\caption{K09-b vulnerable: all pods share the \texttt{default} SA with auto-mounted token.}\label{lst:k09-sharedSA-vuln}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k09-sharedSA-safe.yaml}
\caption{K09-b remediated: per-service \texttt{ServiceAccount} with \texttt{automountServiceAccountToken: false}.}\label{lst:k09-sharedSA-safe}
\end{listing}

% ─── K10 ────────────────────────────────────────────────────────────────────
\subsubsection{K10 --- Inadequate Audit Logging (2022: K05)}
\label{sec:k10}

The Kubernetes API server can emit structured audit events for every API call, but this feature is disabled by default~\cite{k8s-audit-docs}.
Without an audit policy, there is no forensic record of secret access, privilege escalation, or lateral movement.

\paragraph{K10-a: No \texttt{kube-apiserver} audit policy configured (CWE-778, CIS~1.2.16).}\label{sec:k10-noAuditPolicy}
Minikube starts the API server without \texttt{-{}-audit-policy-file}, \texttt{-{}-audit-log-path}, or related flags.
All API operations are silent: an attacker who lists all Secrets, binds themselves to \texttt{cluster-admin}, and exfiltrates data leaves no server-side log~\cite{k8s-audit-docs}.
This entry is cluster-level and was \emph{already present} in the baseline (\texttt{presentPreInjection: true}).
The exploit script (Listing~\ref{lst:k10-auditPolicy-exploit}) confirms the absence by inspecting the running \texttt{kube-apiserver} process arguments.
Safe mode requires copying \texttt{k8s/insecure-defaults/audit-policy-safe.yaml} to the Minikube node and restarting the API server with the audit flags (see \texttt{audit-policy-missing.md}).
Detected by: Trivy \texttt{KCV0019}, KICS rule \texttt{13a49a2e}, Kubescape \texttt{C-0130}~\cite{trivy-k8s,kics-tool,kubescape}.

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{bash}{snippets/k10-auditPolicy-exploit.sh}
\caption{K10-a exploit script: confirms absence of \texttt{-{}-audit-policy-file} on running API server.}\label{lst:k10-auditPolicy-exploit}
\end{listing}

\begin{listing}[ht]
\inputminted[fontsize=\footnotesize,frame=lines]{yaml}{snippets/k10-auditPolicy-safe.yaml}
\caption{K10-a remediated: audit policy capturing secrets access, pod exec, RBAC changes, and auth failures.}\label{lst:k10-auditPolicy-safe}
\end{listing}

% ─── Discussion ──────────────────────────────────────────────────────────────
\subsection{Discussion}
\label{sec:k8s-discussion}

\paragraph{Outdated Kubernetes components (K10-related).}\label{sec:k8s-outdatedComponents}
Zerouali et al.~\cite{zerouali-2023} demonstrate that a significant fraction of Docker images in the wild carry outdated system packages.
The same phenomenon applies to the Kubernetes version itself: Minikube ships with a fixed Kubernetes release that accumulates \textsc{cve}s over time.
MAD~Goat does not inject a specific outdated-components vulnerability because the benchmark's Minikube target version is already configurable via \texttt{minikube start -{}-kubernetes-version}; this axis is a natural extension for future benchmark releases.

\paragraph{Log aggregation gap (K10-related).}\label{sec:k8s-logAggregation}
The K10-a audit policy (Section~\ref{sec:k10-noAuditPolicy}) writes events to a local file on the Minikube node.
A production deployment requires shipping these events to a central SIEM to survive node failure and support cross-cluster correlation.
The benchmark documents this gap in \texttt{k8s/insecure-defaults/audit-policy-missing.md} but defers log-shipping configuration to the operator, as SIEM integration is out of scope for a local benchmark cluster.

\paragraph{Scanner coverage gaps.}\label{sec:k8s-scannerGaps}
Table~\ref{tab:scanner-coverage} reveals that application-layer misconfigurations (K06-a Traefik dashboard, K06-b RabbitMQ management, K09-a Keycloak \texttt{start-dev}) produce no findings in any of the four tested scanners.
This is consistent with findings by Bose et al.~\cite{bose-2023} that a class of ``under-reported'' Kubernetes security defects exists outside the scope of manifest-level static analysis.
These entries require application-aware or behavioural detection methods, such as custom Kyverno rules tuned to the specific service or network-anomaly detection on management-plane ports.

% ─── Summary ─────────────────────────────────────────────────────────────────
\subsection{Summary}
\label{sec:k8s-summary}

This section has extended the MAD~Goat benchmark with 20 toggle-controlled vulnerability entries spanning all ten OWASP Kubernetes Top~10 (2025) categories.
Fifteen entries are implemented via Helm chart boolean flags in \texttt{helm/madgoat} and \texttt{helm/madgoat-infra}; five cluster-level entries are applied via \texttt{kubectl apply -k k8s/insecure-defaults/} and Minikube startup flags.
Each entry ships with a verification script in \texttt{scripts/exploits/}, and a \texttt{Makefile} target (\texttt{make reset-vulnerable} / \texttt{make reset-safe}) provisions the full vulnerable or hardened cluster state.
Two entries (K08) are conceptual-only, reflecting the cloud-specific nature of IMDS and workload-identity attacks.
Table~\ref{tab:scanner-coverage} quantifies the detection gap: of the four scanners evaluated, none detect application-layer authentication defaults (K06, K09-a), underscoring the need for application-aware policy rules as a complement to generic manifest scanning~\cite{bose-2023,rahman-2019}.

\input{tables/scanner-coverage}